---
title: "Précision sur le Blended cours"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Meta-analysis for Data Science and Business Analytics

* 1) Why do we need for standard measure ?

A - to avoid effect size

B - to calculate the mean

C - to calculate standard deviation


```{r}
# Casser le nombre d'individus par classe
table(iris$Species)
df_desequilibre<- iris[1:101 , ]
table(df_desequilibre$Species)

library(dplyr)
df_desequilibre %>% 
        as_tibble() %>% 
        group_by(Species) %>% 
        summarise_if(is.numeric, mean)
```


```{r}
iris %>% 
        as_tibble() %>% 
        group_by(Species) %>% 
        summarise_if(is.numeric, mean)

```

```{r}
# la grandeur (moyenne) des variables et la variance
```



* 2) Suppose that you wanted to test the difference between men's heights in inches and women's height in inches to quantify the effect of gender on height. 

In this case, what are the independent and dependent variables?

A - Gender (independent) and Height (dependent)

B - Gender (dependent) and Height (independent)

C - None of the above


* 3) Suppose that you wanted to test the difference between men's heights in inches women's height in inches to quantify the effect of gender on height. 

Can we use, in this case, an analysis of the variance to explain the difference in the mean height in men and women?


## Business Analytics Foundations - Predictive, Prescriptive, and Experimental Analytics

* Dashboards/reports and Exploratory Analytics belong to :

A - Business Intelligence

B - Business Analytics

C - Belong to both

 
* Machine Learning and Statistical Modeling  belong to :

A - Business Intelligence

B - Business Analytics

C - Belong to both


* How many stages of Business Analytics?

A - 3

B - 4

C - 5

D - 6

 
* To make prediction, we need  a large quantity of data and :

A - Historical data

B- Statistics

C - Machine Learning

D - All of  the above

 
* The goal of Prescriptive Analytics is to explain :

A - How do I take Advantage?

B - What will Happen?

C - Why did it Happen?

 
* The goal of Experimental Analytics is to understand actual performance by testing

 a hypothesis or an alternative
 
 A - True
 
 B - False


```{r}
f <- rnorm(100, 60, 10)
h <- rnorm(100, 70, 10)
library(dplyr)
dd <- tibble("f" = f, "h" = h)

```

```{r}
dd2 <- dd[1:70 , ]

summary(dd2)

library(ggplot2)
dd2 %>% 
        tidyr::pivot_longer(. , cols = c("f", "h")) %>%
        ggplot(. , aes(x = value, color = name)) +
        geom_density()
```
 
 
 
```{r}
dd2 %>% 
        tidyr::pivot_longer(. , cols = c("f", "h")) %>%
        ggplot(. , aes(y = value, color = name)) +
        geom_boxplot()
```
 
 
 
```{r}

# en appliquant la règle de Kévin taille < 65 => Genre = Femme
table(dd2[dd2$f <65 | dd2$h < 65, ] )
sum(dd2$f <65)
sum(dd2$h < 65)

str(dd2)

sum(dd2$f >= 65)
sum(dd2$h >= 65)

```


```{r}
d_exp <- dd[71:100 , ]
# str(dd)

sum(d_exp$f <65)
sum(d_exp$h < 65)
sum(d_exp$f >= 65)
sum(d_exp$h >= 65)

```


##