---
title: "TP2 : Linear Regression"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Basic linear regression

Let $x_1$ and $x_2$ be 2 explanatory variables and $y$ the dependent variable. $\epsilon$ (espilon) the error of  prediction (supposed to be a random variable). We modelise $y = b_0 + b_1x_1 + b_2x_2 + \epsilon$. The question is how to find the best coefficients $b_0$, $b_1$ and $b_2$ to minimize the error of modelisation : $(y - \hat y)^2$ with $\hat y$ the calculated estimation of $y$ by $\hat y = b_0 + b_1x_1 + b_2x_2$.

The function `lm` of R makes it possible to calculate the linear regression of a numerical dependent variable as a function of explanatory variables (here $y = f(x_1, x_2) + \epsilon$).

#### 1 - Use `lm` on a simple example

You can use the help panel (or press F1 with the word selection) to fetch R's help documentation.

Here the formula is reduced to $y = b_0 + b_1x + \epsilon$.

```{r}
x <- c(4, 6, 3, 5, 1, 9)
y <- 2 * x + 1 + rnorm(6, 0, 0.3)
```


#### 2 - With a data frame

```{r}
fr <- data.frame(x = c(4, 6, 3, 5, 1, 9), y = c(8, 11, 5, 10, 3, 17), z = c("a", "a", "b", "a", "b", "a"))
```


#### 3 - Draw the graph with the line using plot function

```{r}

```


#### 4 - Interpret the object, of class "lm"

```{r}

```

#### 5 - Give the confidence interval on the coefficients, with alpha = 0.95

```{r}

```


#### 6 - Prediction of values
The predict function returns a vector of the predicted values (names of the explanatory variables must be the same in the data frame).

```{r}

```


#### 7 - Plot the confidence intervals (of the predicted values)

```{r}

```


#### 8 - Goodness of the model

Calculate the mean square error and comment on its meaning.

```{r}

```

#### 9 - Plot the coefficients interval of confidence

```{r}

```


## Using the diamond data

A common method of machine learning is to use a train sample to train the algorithm and then to evaluate the goodness of the model on a test sample.

Use the following train and test sample from the `diamonds` dataset to predict the `price` of a diamond in function of its `carat` and `clarity` value.


```{r}
diamonds <- ggplot2::diamonds
# str(diamonds)
set.seed(123)
sample_index <- sample(x = 1:nrow(diamonds), size = floor(0.8*nrow(diamonds)), replace = FALSE)
train <- diamonds[sample_index , ]
test <- diamonds[-sample_index , ]
# str(train)
# str(test)
```

